Helmet Foam Case Study Report

### 2024-05-07 CS2 Final Draft.qmd
Add interpretations

We reformatted the data from wide to long to ensure each observation represented one measurement, not one foam sample. Measurements taken beyond 10,000 cycles were removed to maintain consistency across observations, and because Moira identified this as the most relevant time frame.

@@

ggplot(foam_Samples, aes(x=freq)) +
  geom_histogram(bins=10)
```

The histogram in @fig-1 shows the distribution of compression frequencies among the foam samples revealing that certain compression frequencies were tested more frequently than others. These non-uniformities in sample distribution might impact the reliability of conclusions drawn from the analysis.

@@

combined
```
In order to see how stiffnessFrac and dampingFrac change over time for all foams, we plotted a scatter plot for each vs cycle. The dot colors in @fig-2 represent which foam sample they were measured from. As cycle number increases, the stiffnessFrac and dampingFrac values not only change, but spread further apart. StiffnessFrac and dampingFrac both appear to have a relationship with cycle, but that relationship looks different for each foam sample.

@@

We also see an effect from porosity. For chemical index 79, low porosity foam trends downward more often than high porosity foam, and for chemical index 100, low porosity foam trends upward more often than high porosity foam. So not only does porosity affect log(stiffnessFrac), but it effects it different ways for different porosities. This suggests that a porosity:chemIndex interaction term would improve our model.

@@

Porosity also appears to have some effect. Trends in log(dampingFrac) for chemical index 121 are more similar to each other for low porosity than for high porosity, and trends in log(dampingFrac) for chemical index 79 are bit more similar to each other for high porosity than for low porosity. This suggests a porosity:chemIndex interaction term would improve our damping model, but this is still less clear than it was for stiffness.

@@

Based on the research questions, we developed two separate models: one for stiffness fraction and the other for damping fraction. Due to the experimental design, and the fact that our observations are not independent of each other, we used a mixed model format for our predictions. Log(cycle) was treated as a level-one variable because it changes across individual measurements, while all other variables were treated as level-two variables because they only vary across foam samples. In a mixed model, level-two variables impact the model intercept, and level-one variables interact with the level-two variables to effect the model slope.

@@

@@

All samples have a stiffness fraction of 1 at cycle zero. A mixed model can create individual intercepts for each sample, so for this model, we used samples starting at the first cycle measurement after cycle zero. Thus, $\alpha$ terms in the model  explain the impact of the covariates right after compressions begin, and the $\beta$ terms explain the impact of the covariates as time goes on.

@@

In Table 5, we see high significance (aka p-value < 0.05) for most of our coefficients. Some variables are not significant on their own, their interactions with log(cycle) are. For example, chemical index does not appear to have a significant relationship with the response on it's own, but one level of chemical index has a significant relationship when interacted with log(cycle). The interaction terms explain how log(cycle) effects stiffness fraction differently under different conditions. From this summary, amplitude appears to have the largest positive impact on initial stiffness fraction, and the largest negative impact as time progresses.

@@

A similar concern comes from foam types and testing conditions that were used less often, such as high frequency. The reliability of our models is already limited, and for these cases they are even less reliable. Additionally, if one variety of foam was never tested at high frequency, then the model's claims about that foam at high frequency are based on nothing, and should not be trusted at all. So not all information this model gives on what effects stiffness and damping in foam is equally trustworthy.

@@

ggplot(foam_Samples, aes(x=freq)) +
  geom_histogram(bins=10)

>>>>

We reformatted the data from wide to long to ensure each observation represented one measurement, not one foam sample. Measurements taken beyond 10,000 cycles were removed to maintain consistency across observations, and because Moira identified this as the most relevant time frame. The first measurements for stiffnessFrac and dampingFrac were removed as well, because they were all equal to 1 and did not convey useful information.

@@

ggplot(foam_Samples, aes(x=amp)) +
  geom_histogram(bins=10)
```

The histogram in @fig-1 shows the distribution of compression amplitudes among the foam samples, revealing that certain amplitudes were tested much less frequently than others. These non-uniformities in sample distribution might impact the reliability of conclusions drawn from the analysis.

@@

combined
```

In order to see how stiffnessFrac and dampingFrac change over time for all foams, we plotted a scatter plot for each vs cycle. The dot colors in @fig-2 represent which foam sample they were measured from. As cycle number increases, the stiffnessFrac and dampingFrac values not only change, but spread further apart. StiffnessFrac and dampingFrac both appear to have a relationship with cycle, but that relationship looks different for each foam sample.

@@

We also see an effect from porosity. For chemical index 79, low porosity foam trends downward more often than high porosity foam, and for chemical index 100, low porosity foam trends upward more often than high porosity foam. So not only does porosity affect log(stiffnessFrac), but it effects it different ways for different chemical indices This suggests that a porosity:chemIndex interaction term might improve our model.

@@

Porosity also appears to have some effect. Trends in log(dampingFrac) for chemical index 121 are more similar to each other for low porosity than for high porosity, and trends in log(dampingFrac) for chemical index 79 are bit more similar to each other for high porosity than for low porosity. This suggests a porosity:chemIndex interaction term could improve our damping model, but this is still less clear than it was for stiffness.

@@

Based on the research questions, we developed two separate models: one for stiffness fraction and the other for damping fraction. Due to the experimental design, and the fact that measurements from the same sample are not independent of each other, we used a mixed model format for our predictions. Log(cycle) was treated as a level-one variable because it changes across individual measurements, while all other variables were treated as level-two variables because they only vary across foam samples. In a mixed model, level-two variables impact the model intercept, and level-one variables interact with the level-two variables to effect the model slope. This accounts for each sample appeared to react to log(cycle) differently.

@@

Despite the expectations we got from figures 4 and 5, we could not determine that an interaction between chemical index and porosity was as significant as the predictors listed in Table 4, so it was not included.

@@

The $\alpha$ terms in the model explain the impact of the covariates right after compressions begin (intercepts), and the $\beta$ terms explain the impact of the covariates as time goes on (slopes).

@@

In Table 5, we see high significance (aka p-value < 0.05) for most of our coefficients. Some variables are not significant on their own, their interactions with log(cycle) are. For example, chemical index does not appear to have a significant relationship with the response on it's own, but one level of chemical index has a significant relationship when interacted with log(cycle). The interaction terms explain how log(cycle) effects stiffness fraction differently under different conditions.

In this summary, log(cycle) has a positive relationship with stiffnessFrac, but that is only when frequency and amplitude are zero. The interactions of freq and amp with log(cycle) have negative coefficients, showing us that stiffness decreases more when compressions are faster and larger. Chemical index levels 2 and 3 (aka 100 and 121) interacted with log(cycle) also have negative coefficients, indicating that these foams lose stiffness faster than foam with index 79.

@@

A similar concern comes from foam types and testing conditions that were used less often, such as high amplitude The reliability of our models is already limited, and for these cases they are even less reliable. Additionally, if one variety of foam was never tested at high amplitude, then the model's claims about that foam at high frequency are based on nothing, and should not be trusted at all. So not all information this model gives on what effects stiffness and damping in foam is equally trustworthy.

@@

ggplot(foam_Samples, aes(x=amp)) +
  geom_histogram(bins=10)
